Modelling $q_{95}$ Windows for the Suppression of Edge Localized Modes by Resonant Magnetic Perturbations in the DIII-D Tokamak: Relaxation of the No-Slip Constraint

### 2020-09-07 Papers/Paper2/Paper.tex
Phase version 2.3
Added PHASE_LIN flag

\section{Introduction}

\section{Cylindrical RMP Response Model}\label{s2}
\subsection{Plasma Equilibrium}
Consider a large aspect-ratio, low-$\beta$, tokamak plasma whose magnetic flux surfaces map out (almost)
concentric circles in the poloidal plane.
Such a plasma is well approximated as a periodic cylinder.  Suppose that the minor radius of the plasma
is $a$. Standard cylindrical coordinates ($r$, $\theta$, $z$) are adopted. The system is assumed to
be periodic in the $z$-direction, with periodicity length $2\pi\,R_0$, where $R_0\gg a$ is the simulated
plasma major radius. It is convenient to define the simulated toroidal angle $\phi=z/R_0$.

The equilibrium magnetic field is written ${\bf B} = [0,B_\theta(r), B_\phi]$. The associated equilibrium
plasma current density takes the form ${\bf j} =[0,0,j_\phi(r)]$, where $\mu_0\,j_\phi= (1/r)\,d(r\,B_\theta)/dr$. The
so-called {\em safety-factor}, $q(r)=r\,B_\phi/(R_0\,B_\theta)$, parameterizes the helical pitches of equilibrium magnetic
field-lines.

\subsection{Plasma Response to RMP}
Consider the response of the plasma to an externally generated, static, helical, resonant magnetic perturbation (RMP). 
Suppose that the RMP has $m$ periods in the poloidal direction, and $n$ periods in the toroidal direction.
It is convenient to express the perturbed magnetic field and the perturbed plasma current density in
terms of a perturbed poloidal magnetic flux, $\psi(r,\theta,\phi,t)$. In fact, $\delta{\bf B} = \nabla\psi\times \nabla z$, 
and $\mu_0\,\delta {\bf j} = - \nabla^2\psi\,\nabla z$,
where $\psi(r,\theta,\phi,t) = \hat{\psi}(r,t)\,\exp[{\rm i}\,(m\,\theta-n\,\phi)]$, and $\hat{\psi}$ is real. This particular representation
is valid provided that $m/n\gg a/R_0$.\cite{rfa}

The response of the plasma to the RMP is governed by the equations of marginally stable, ideal-MHD everywhere in the
plasma, apart from a relatively narrow (in $r$) region in the vicinity of the so-called {\em rational surface}, minor radius $r_s$, 
at which $q(r_s)=m/n$.\cite{rfa}

It is convenient to parameterize the RMP in terms of the so-called {\em vacuum magnetic flux}, ${\mit\Psi}_v(t) = |{\mit\Psi}_v|\,{\rm e}^{-{\rm i}\,\varphi_v}$,
which is defined to be the value of $\hat{\psi}(r,t)$ at radius $r_s$ in the presence of the RMP, but in the absence of plasma. Here,
$\varphi_v$ is the helical phase of the RMP. Likewise, the response of the plasma in the vicinity of the rational
surface is parameterized in terms of the so-called {\em reconnected magnetic flux}, ${\mit\Psi}_s(t)=|{\mit\Psi}_s|\,{\rm e}^{\,{\rm i}\,\varphi_s}$, 
which is the actual value of $\hat{\psi}(r,t)$ at radius $r_s$. Here, $\varphi_s$ is the helical phase of
the reconnected magnetic flux.

The intrinsic stability of the $m$, $n$ tearing mode is governed by the dimensionless parameter
$E_{ss}= [d\hat{\psi}_s/d\ln r]_{r_{s-}}^{r_{s+}}$, where $\hat{\psi}_s(r)$
is a solution of the marginally stable, ideal-MHD equations, for the case of an $m$, $n$ helical
perturbation, that satisfies physical boundary conditions at $r=0$ and $r=a$, in the absence of the
RMP, and is such that $\hat{\psi}_s(r_s)=1$.\cite{fkr,am1} According to resistive-MHD theory,\cite{fkr,ruth}
if $E_{ss}>0$ then the $m$, $n$ tearing mode spontaneously reconnects magnetic flux at the rational
surface to form a helical magnetic island chain. In this paper, it is assumed that $E_{ss}<0$, so that the
$m$, $n$ tearing mode is intrinsically stable. In this case, any magnetic reconnection that takes place
at the rational surface is due solely to the action of the RMP.

The ideal-MHD response of the plasma to the RMP is governed by the dimensionless parameter
$E_{sv}= [d\hat{\psi}_v/d\ln r]_{r_{s-}}^{r_{s+}}$, where $\hat{\psi}_v(r)$
is a solution of the marginally stable, ideal-MHD equations, for the case of an $m$, $n$ helical
perturbation, that satisfies physical boundary conditions at $r=0$ and $r=a$, in the presence of the
RMP, and is such that $\hat{\psi}_v(r_s)=0$.\cite{am1}

\subsection{Linear Response Regime}
In the linear response regime, the reconnected magnetic flux induced by the RMP at the
rational surface is governed by\,\cite{rfa,rfb,rfc}
\begin{equation}\label{e1}
\frac{\delta_s}{r_s}\,\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v,
\end{equation}
where $\delta_s$ is the linear layer width, $\tau_R=\mu_0\,r_s^{\,2}\,\sigma(r_s)$ the global resistive diffusion timescale,
 and
\begin{equation}
\omega_s (t)=m\,{\mit\Omega}_\theta(r_s,t)-n\,{\mit\Omega}_\phi(r_s,t).
\end{equation}
Here, $\sigma(r)$, ${\mit\Omega}_\theta(r,t)$, and ${\mit\Omega}_\phi(r,t)$ are the 
plasma conductivity, poloidal angular velocity, and toroidal angular velocity profiles, respectively. 
Note that, in this paper, we are assuming that the constant-$\psi$ approximation\,\cite{fkr} holds at the rational surface. (It is demonstrated in
Ref.~\onlinecite{rf1} that the appropriate linear response regime for an RMP resonant in the pedestal of a typical DIII-D H-mode discharge
is the so-called  {\em first semi-collisional regime}, which is indeed a constant-$\psi$ response regime. Moreover, if the appropriate linear response regime
is constant-$\psi$ then so is the appropriate nonlinear response regime, because the constant-$\psi$ constraint is
easier to satisfy in the nonlinear regime than in the linear regime.\cite{rfb,rffig})

Equation~(\ref{e1}) can be rewritten as an {\em amplitude evolution equation},
\begin{equation}\label{e3}
\frac{\delta_s}{r_s}\,\tau_R\,\frac{d|{\mit\Psi}_s|}{dt}= E_{ss}\,|{\mit\Psi}_s| + E_{sv}\,|{\mit\Psi}_v|\,\cos(\varphi_s-\varphi_v),
\end{equation}
combined with a {\em phase evolution equation}, 
\begin{equation}\label{e4}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{\delta_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of Eq.~(\ref{e4}) is termed the {\em slip frequency}, and is the difference between the helical phase velocity of
the reconnected magnetic flux and that expected on the  assumption that the flux is convected by the plasma flow at the
rational surface (i.e., $d\varphi_s/dt=\omega_s$). In general, the slip frequency is non-zero because the plasma in the vicinity of the rational surface is capable of diffusing resistively 
through the linear layer structure.

\subsection{Nonlinear Response Regime}
In the nonlinear response regime, Eq.~(\ref{e3}) generalizes to the {\em Rutherford island width evolution equation}:\,\cite{ruth,ruth1}
\begin{equation}\label{e5}
\frac{({\cal I}/2)\,W_s}{r_s}\,\tau_R\,\frac{d|{\mit\Psi}_s|}{dt} = E_{ss}\,|{\mit\Psi}_s| + E_{sv}\,|{\mit\Psi}_v|\,\cos(\varphi_s-\varphi_v),
\end{equation}
where
\begin{equation}
W_s = 4\left(\frac{|{\mit\Psi}_s|}{s_s\,r_s\,B_\theta(r_s)}\right)^{1/2} r_s
\end{equation}
is the full radial width of the magnetic island chain, ${\cal I} = 0.8227$, $s(r) = d\ln q/d\ln r$, and $s_s=s(r_s)$. The
nonlinear regime holds when $W_s\gg \delta_s$, whereas the linear regime holds when $\delta_s\ll W_s$.

Equation~(\ref{e5}) is usually coupled with the so-called {\em no-slip constraint}:\,\cite{rfa}
\begin{equation}\label{e6}
\frac{d\varphi_s}{dt} = \omega_s.
\end{equation}
The reasoning behind the imposition of this constraint is that the plasma is trapped inside the
magnetic separatrix of the magnetic island chain that forms in the vicinity of the rational surface,
which forces the chain to co-rotate with the local plasma flow. However, in reality, the plasma
is able to diffuse resistively across the separatrix to some extent. Hence, by analogy with Eqs.~(\ref{e3}), (\ref{e4}), and (\ref{e5}), and 
in accordance with Refs.~\onlinecite{slip1} and \onlinecite{slip2}, in this paper we shall modify Eq.~(\ref{e6}) such that
\begin{equation}\label{e7}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{({\cal I}/2)\,W_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of the previous equation is the {\em nonlinear slip frequency}. In general,
we would expect this frequency to be relatively small (because it is generally the case that $|\omega_s|\,\tau_R\,(W_s/r_s)\gg 1$
for fully-developed magnetic islands in high-temperature tokamak plasmas\,\cite{rf1,slip2}). Nevertheless, the
nonlinear slip frequency may be non-negligible for developing island chains, which is why we are including it in our analysis.
(Note that the nonlinear slip frequency does not appear in rigorous island calculations, such as
that described in Ref.~\onlinecite{rfisland}, and references therein, because such calculations depend crucially on
complicated hierarchical ordering assumptions, according to which the nonlinear slip frequency is negligible.)

\subsection{Composite Model}
We can combine Eqs.~(\ref{e3}), (\ref{e4}), (\ref{e5}), and (\ref{e7}) to give the following
composite resonant response model that interpolates between the linear and the nonlinear
regimes:
\begin{align}\label{e9}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\,\frac{d|{\mit\Psi}_s|}{dt} &= E_{ss}\,|{\mit\Psi}_s| + E_{sv}\,|{\mit\Psi}_v|\,\cos(\varphi_s-\varphi_v),\\[0.5ex]
\frac{d\varphi_s}{dt}& = \omega_s - \frac{E_{sv}}{\tau_R}\left(\frac{r_s}{({\cal I}/2)\,W_s+\delta_s}\right)\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).\label{e10}
\end{align}
If we define $X_s = |{\mit\Psi}_s|\,\cos\varphi_s$, $Y_s = |{\mit\Psi}_s|\,\sin\varphi_s$, $X_v = |{\mit\Psi}_v|\,\cos\varphi_v$, and $Y_v = |{\mit\Psi}_v|\,\sin\varphi_v$ 
then the previous two equations are more conveniently written in the nonsingular (when $|{\mit\Psi}_s|=0$) forms:\,\cite{slip1,slip2,slip3}
\begin{align}\label{e11}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dX_s}{dt}+\omega_s\,Y_s\right)&= E_{ss}\,X_s + E_{sv}\,X_v,\\[0.5ex]
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dY_s}{dt}-\omega_s\,X_s\right)& = E_{ss}\,Y_s + E_{sv}\,Y_v,\label{e12}
\end{align}
where
\begin{equation}
W_s = 4\left(\frac{\sqrt{X_s^{\,2}+Y_s^{\,2}}}{s_s\,r_s\,B_\theta(r_s)}\right)^{1/2} r_s.
\end{equation}
With the benefit of hindsight, Eqs.~(\ref{e11}) and (\ref{e12}) could have been derived more directly from the
following heuristic generalization of Eq.~(\ref{e1}):\,\cite{slip2}
\begin{equation}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v.
\end{equation}

\section{Modified Toroidal RMP Response Model}
Our toroidal model of the response of a tokamak plasma to a static, externally generated, RMP is described in
detail in Ref.~\onlinecite{rftor}. In the light of the composite model developed in Sect.~\ref{s2}, we shall
modify the resonant plasma response component of our  toroidal response model; this component is described in Appendix~D of Ref.~\onlinecite{rftor}.

Equations~(D1) and (D2) in Ref.~\onlinecite{rftor} are modified such that 
\begin{align}\label{e15}
\left(\hat{W}_k + \hat{\delta}_k\right){\cal S}_k\,\frac{d\hat{\mit\Psi}_k}{d\hat{t}} &= \sum_{k'=1,K}\hat{E}_{kk'}\hat{\mit\Psi}_{k'}\,\cos(\varphi_k-\varphi_{k'}-\xi_{kk'})
+\hat{E}_{kk}\,\hat{\chi}_k\,\cos(\varphi_k-\zeta_k),\\[0.5ex]
{\cal S}_k&=\frac{\tau_R(\hat{r}_k)}{\tau_A},
\end{align}
respectively. Here, 
\begin{align}
\hat{W}_k &=\frac{2\,{\cal I}}{\hat{a}\,\hat{r}_k}\left(\frac{q}{g\,s}\right)_{\hat{r}_k} |\hat{\mit\Psi}_k|^{\,1/2},\\[0.5ex]
\hat{\delta}_k &= \frac{\delta_{\rm SC}(\hat{r}_k)}{R_0\,\hat{a}\,\hat{r}_k},
\end{align}
where $\delta_{\rm SC}(\hat{r})$ is the semi-collisional linear layer width specified in Ref.~\onlinecite{rf1}.

@@

Equation~(D6) of Ref.~\onlinecite{rftor} is modified such that
\begin{align}\label{e19}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{d\varphi_k}{d\hat{t}}-\hat{\varpi}_k\right)&=
- \sum_{k'=1,K}\hat{E}_{kk'}\,\hat{\mit\Psi}_{k'}\,\sin(\varphi_k-\varphi_{k'}-\xi_{kk'})-\hat{E}_{kk}\,\hat{\chi}_k\,\sin(\varphi_k-\zeta_k),
\end{align}
where
\begin{equation}
\hat{\varpi}_k= m_k\,\hat{\mit\Omega}_\theta(\hat{r}_k,\hat{t})-n\,\hat{\mit\Omega}_\phi(\hat{r}_k,\hat{t}),
\end{equation}
and $\hat{\mit\Omega}_\theta= {\mit\Omega}_\theta\,\tau_A$, $\hat{\mit\Omega}_\phi= {\mit\Omega}_\phi\,\tau_A$. Equation~(\ref{e19}) represents
a major improvement to Eq.~(D6) that takes into  account the finite slip frequencies of the magnetic island chains induced at
the various rational surfaces in the plasma.

>>>>

\section{Introduction}
In the context of tokamak plasmas, a resonant magnetic perturbation (RMP) refers to an externally
generated, (usually) static, helical magnetic perturbation that resonates at
one or more rational surfaces within the plasma. Resonant magnetic perturbations have been successfully
employed to compensate error fields (i.e., accidentally produced RMP), and, thereby,  to prevent the formation of
locked modes in a wide variety of different types of tokamak discharge. \cite{error1,error2,error3} RMPs have also been used to either control or
mitigate edge-localized-modes (ELMs) in H-mode tokamak discharges.\cite{elm1,elm2,elm3,elm4,elm5,elm6}

By far the most efficient manner in which to model the response of a tokamak plasma to an RMP is
to employ an ``asymptotic matching'' approach.\cite{fkr,rfa, am1,am2,am3} 
According to such an approach, the response of the plasma to the applied RMP is governed by a combination of flux-freezing and
perturbed force balance---this combination is usually referred to as ``marginally-stable ideal-magnetohydrodynamics (MHD)", which is a misnomer because MHD {\em per se}\/ plays no
role---everywhere in the plasma apart from a number of relatively narrow (in the radial direction) regions in which the applied
perturbation resonates with the equilibrium magnetic field. Magnetic reconnection can take place within the resonant regions to
produce relatively thin magnetic island chains. Within the resonant regions, the plasma response is governed by  linear or nonlinear 
 two-fluid resistive-MHD, depending on whether the induced magnetic island widths are smaller or larger than the corresponding linear layer widths. 
 Thus, when employing the asymptotic matching approach, the equations of marginally-stable ideal-MHD
are solved in the so-called ``outer region'' that comprises most of the plasma (and the surrounding vacuum), the equations of 
linear/nonlinear two-fluid resistive-MHD are solved in the various resonant layers that constitute the so-called ``inner region'', and the two
sets of solutions are then asymptotically matched to one another.

In the nonlinear resonant response regime, the evolution of the island width in a given resonant layer is governed by the familiar Rutherford island width evolution equation.\cite{ruth,ruth1}
On the other hand, the phase evolution of the island chain is conventionally governed by the well-known no-slip constraint.\cite{rfa} According to this constraint,  plasma is trapped inside the
magnetic separatrix of the magnetic island chain,
which forces the chain to co-rotate with the local plasma flow. In reality, however, the plasma
is able to diffuse resistively across the separatrix to some extent.\cite{slip1,slip2}

The primary aim of this paper is to investigate 
what happens when the no-slip constraint is relaxed in a recently-developed model of  RMP-induced
suppression of ELMs in H-mode tokamak discharges.\cite{rftor} The secondary aim is to examine what happens when an improvement is made to the manner in which the resonant
plasma response component of the model interpolates between the linear and the nonlinear regime.

This paper is organized as follows. In Sect.~\ref{s2}, we develop a cylindrical version of
our new resonant plasma response model. In Sect.~\ref{s3},  we use this cylindrical version as a guide for producing a toroidal
version of the model.

\section{Cylindrical Resonant Response Model}\label{s2}
\subsection{Plasma Equilibrium}
Consider a large aspect-ratio, low-$\beta$, tokamak plasma whose magnetic flux surfaces map out (almost)
concentric circles in the poloidal plane.
Such a plasma is well approximated as a periodic cylinder.  Suppose that the minor radius of the plasma
is $a$. Standard cylindrical coordinates ($r$, $\theta$, $z$) are adopted. The system is assumed to
be periodic in the $z$-direction, with periodicity length $2\pi\,R_0$, where $R_0\gg a$ is the simulated
plasma major radius. It is convenient to define the simulated toroidal angle $\phi=z/R_0$.

The equilibrium magnetic field is written ${\bf B} = [0,B_\theta(r), B_\phi]$. The associated equilibrium
plasma current density takes the form ${\bf j} =[0,0,j_\phi(r)]$, where $\mu_0\,j_\phi= (1/r)\,d(r\,B_\theta)/dr$. The
so-called ``safety-factor'', $q(r)=r\,B_\phi/(R_0\,B_\theta)$, parameterizes the helical pitches of equilibrium magnetic
field-lines.

\subsection{Plasma Response to RMP}
Consider the response of the plasma to an externally generated, static, helical, RMP. 
Suppose that the RMP has $m$ periods in the poloidal direction, and $n$ periods in the toroidal direction.
It is convenient to express the perturbed magnetic field and the perturbed plasma current density in
terms of a perturbed poloidal magnetic flux, $\psi(r,\theta,\phi,t)$. In fact, $\delta{\bf B} = \nabla\psi\times \nabla z$, 
and $\mu_0\,\delta {\bf j} = - \nabla^2\psi\,\nabla z$,
where $\psi(r,\theta,\phi,t) = \hat{\psi}(r,t)\,\exp[{\rm i}\,(m\,\theta-n\,\phi)]$, and $\hat{\psi}$ is real. This particular representation
is valid provided that $m/n\gg a/R_0$.\cite{rfa}

The response of the plasma to the RMP is governed by the equations of marginally-stable, ideal-MHD everywhere in the
plasma, apart from a relatively narrow (in $r$) region in the vicinity of the so-called  ``rational surface'', minor radius $r_s$, 
at which $q(r_s)=m/n$.\cite{rfa}

It is convenient to parameterize the RMP in terms of the so-called ``vacuum magnetic flux'', ${\mit\Psi}_v(t) = |{\mit\Psi}_v|\,{\rm e}^{-{\rm i}\,\varphi_v}$,
which is defined to be the value of $\hat{\psi}(r,t)$ at radius $r_s$ in the presence of the RMP, but in the absence of plasma. Here,
$\varphi_v$ is the helical phase of the RMP. Likewise, the response of the plasma in the vicinity of the rational
surface is parameterized in terms of the so-called ``reconnected magnetic flux'', ${\mit\Psi}_s(t)=|{\mit\Psi}_s|\,{\rm e}^{\,{\rm i}\,\varphi_s}$, 
which is the actual value of $\hat{\psi}(r,t)$ at radius $r_s$. Here, $\varphi_s$ is the helical phase of
the reconnected magnetic flux.

The intrinsic stability of the $m$, $n$ tearing mode is governed by the dimensionless parameter
$E_{ss}= [d\hat{\psi}_s/d\ln r]_{r_{s-}}^{r_{s+}}$, where $\hat{\psi}_s(r)$
is a solution of the marginally-stable, ideal-MHD equations, for the case of an $m$, $n$ helical
perturbation, that satisfies physical boundary conditions at $r=0$ and $r=a$, in the absence of the
RMP, and is such that $\hat{\psi}_s(r_s)=1$.\cite{fkr,am1} According to resistive-MHD theory,\cite{fkr,ruth}
if $E_{ss}>0$ then the $m$, $n$ tearing mode spontaneously reconnects magnetic flux at the rational
surface to form a helical magnetic island chain. In this paper, it is assumed that $E_{ss}<0$, so that the
$m$, $n$ tearing mode is intrinsically stable. In this case, any magnetic reconnection that takes place
at the rational surface is due solely to the action of the RMP.

The ideal-MHD response of the plasma to the RMP is governed by the dimensionless parameter
$E_{sv}= [d\hat{\psi}_v/d\ln r]_{r_{s-}}^{r_{s+}}$, where $\hat{\psi}_v(r)$
is a solution of the marginally-stable, ideal-MHD equations, for the case of an $m$, $n$ helical
perturbation, that satisfies physical boundary conditions at $r=0$ and $r=a$, in the presence of the
RMP, and is such that $\hat{\psi}_v(r_s)=0$.\cite{am1}

\subsection{Linear Response Regime}
In the linear response regime, the reconnected magnetic flux induced by the RMP at the
rational surface is governed by\,\cite{rfa,rfb,rfc}
\begin{equation}\label{e1}
\frac{\delta_s}{r_s}\,\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v,
\end{equation}
where $\delta_s$ is the linear layer width, $\tau_R=\mu_0\,r_s^{\,2}\,\sigma(r_s)$ the global resistive diffusion timescale,
 and
\begin{equation}
\omega_s (t)=m\,{\mit\Omega}_\theta(r_s,t)-n\,{\mit\Omega}_\phi(r_s,t).
\end{equation}
Here, $\sigma(r)$, ${\mit\Omega}_\theta(r,t)$, and ${\mit\Omega}_\phi(r,t)$ are the 
plasma conductivity, poloidal angular velocity, and toroidal angular velocity profiles, respectively. 
Note that, in this paper, we are assuming that the constant-$\psi$ approximation\,\cite{fkr} holds at the rational surface. (It is demonstrated in
Ref.~\onlinecite{rf1} that the appropriate linear response regime for an RMP resonant in the pedestal of a typical DIII-D H-mode discharge
is the so-called  ``first semi-collisional regime'', which is indeed a constant-$\psi$ response regime. Moreover, if the appropriate linear response regime
is constant-$\psi$ then so is the appropriate nonlinear response regime, because the constant-$\psi$ constraint is
easier to satisfy in the nonlinear regime than in the linear regime.\cite{rfb,rffig})

Equation~(\ref{e1}) can be rewritten as an ``amplitude evolution equation'',
\begin{equation}\label{e3}
\frac{\delta_s}{r_s}\,\tau_R\,\frac{d|{\mit\Psi}_s|}{dt}= E_{ss}\,|{\mit\Psi}_s| + E_{sv}\,|{\mit\Psi}_v|\,\cos(\varphi_s-\varphi_v),
\end{equation}
combined with a ``phase evolution equation'', 
\begin{equation}\label{e4}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{\delta_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of Eq.~(\ref{e4}) is termed the ``slip-frequency'', and is the difference between the helical phase velocity of
the reconnected magnetic flux and that expected on the  assumption that the flux is convected by the plasma flow at the
rational surface (i.e., $d\varphi_s/dt=\omega_s$). In general, the slip-frequency is non-zero because the plasma in the vicinity of the rational surface is capable of diffusing resistively 
through the linear layer structure.

\subsection{Nonlinear Response Regime}
In the nonlinear response regime, Eq.~(\ref{e3}) generalizes to the ``Rutherford island width evolution equation'':\,\cite{ruth,ruth1}
\begin{equation}\label{e5}
\frac{({\cal I}/2)\,W_s}{r_s}\,\tau_R\,\frac{d|{\mit\Psi}_s|}{dt} = E_{ss}\,|{\mit\Psi}_s| + E_{sv}\,|{\mit\Psi}_v|\,\cos(\varphi_s-\varphi_v),
\end{equation}
where
\begin{equation}
W_s = 4\left(\frac{|{\mit\Psi}_s|}{s_s\,r_s\,B_\theta(r_s)}\right)^{1/2} r_s
\end{equation}
is the full radial width of the magnetic island chain, ${\cal I} = 0.8227$, $s(r) = d\ln q/d\ln r$, and $s_s=s(r_s)$. The
nonlinear regime holds when $W_s\gg \delta_s$, whereas the linear regime holds when $\delta_s\ll W_s$.

Equation~(\ref{e5}) is usually coupled with the so-called ``no-slip constraint'':\,\cite{rfa}
\begin{equation}\label{e6}
\frac{d\varphi_s}{dt} = \omega_s.
\end{equation}
The reasoning behind the imposition of this constraint is that the plasma is trapped inside the
magnetic separatrix of the magnetic island chain that forms in the vicinity of the rational surface,
which forces the chain to co-rotate with the local plasma flow. However, in reality, the plasma
is able to diffuse resistively across the separatrix to some extent. Hence, by analogy with Eqs.~(\ref{e3}), (\ref{e4}), and (\ref{e5}), and 
in accordance with Refs.~\onlinecite{slip1} and \onlinecite{slip2}, in this paper we shall modify Eq.~(\ref{e6}) such that
\begin{equation}\label{e7}
\frac{d\varphi_s}{dt} = \omega_s - \frac{E_{sv}}{\tau_R}\,\frac{r_s}{({\cal I}/2)\,W_s}\,\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).
\end{equation}
The final term on the right-hand side of the previous equation is the ``nonlinear slip-frequency''. In general,
we would expect this frequency to be relatively small (because it is generally the case that $|\omega_s|\,\tau_R\,(W_s/r_s)\gg 1$
for fully-developed magnetic islands in high-temperature tokamak plasmas\,\cite{rf1,slip2}). Nevertheless, the
nonlinear slip-frequency may be non-negligible for developing island chains, which is why we are including it in our analysis.
(Note that the nonlinear slip-frequency does not appear in rigorous island calculations, such as
that described in Ref.~\onlinecite{rfisland}, and references therein, because such calculations depend crucially on
complicated hierarchical ordering assumptions, according to which the nonlinear slip-frequency is negligible.)

\subsection{Composite Model}
We can combine Eqs.~(\ref{e3}), (\ref{e4}), (\ref{e5}), and (\ref{e7}) to give the following
composite resonant response model that interpolates between the linear and the nonlinear
regimes:
\begin{align}\label{e9}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\,\frac{d|{\mit\Psi}_s|}{dt} &= E_{ss}\,|{\mit\Psi}_s| + E_{sv}\,|{\mit\Psi}_v|\,\cos(\varphi_s-\varphi_v),\\[0.5ex]
\frac{d\varphi_s}{dt}& = \omega_s - \frac{E_{sv}}{\tau_R}\left(\frac{r_s}{({\cal I}/2)\,W_s+\delta_s}\right)\frac{|{\mit\Psi}_v|}{|{\mit\Psi}_s|}\,\sin(\varphi_s-\varphi_v).\label{e10}
\end{align}
If we define $X_s = |{\mit\Psi}_s|\,\cos\varphi_s$, $Y_s = |{\mit\Psi}_s|\,\sin\varphi_s$, $X_v = |{\mit\Psi}_v|\,\cos\varphi_v$, and $Y_v = |{\mit\Psi}_v|\,\sin\varphi_v$ 
then the previous two equations are more conveniently written in the nonsingular (when $|{\mit\Psi}_s|=0$) forms:\,\cite{slip1,slip2,slip3}
\begin{align}\label{e11}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dX_s}{dt}+\omega_s\,Y_s\right)&= E_{ss}\,X_s + E_{sv}\,X_v,\\[0.5ex]
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{dY_s}{dt}-\omega_s\,X_s\right)& = E_{ss}\,Y_s + E_{sv}\,Y_v,\label{e12}
\end{align}
where
\begin{equation}\label{e23x}
W_s = 4\left(\frac{\sqrt{X_s^{\,2}+Y_s^{\,2}}}{s_s\,r_s\,B_\theta(r_s)}\right)^{1/2} r_s.
\end{equation}
With the benefit of hindsight, Eqs.~(\ref{e11}) and (\ref{e12}) could have been derived more directly from the
following heuristic generalization of Eq.~(\ref{e1}):\,\cite{slip2}
\begin{equation}
\left(\frac{({\cal I}/2)\,W_s+\delta_s}{r_s}\right)\tau_R\left(\frac{d}{dt}+{\rm i}\,\omega_s\right){\mit\Psi}_s= E_{ss}\,{\mit\Psi}_s+E_{sv}\,{\mit\Psi}_v.
\end{equation}

\section{Modified Toroidal Resonant Response Model}\label{s3}
Our toroidal model of the response of a tokamak plasma to a static, externally generated, RMP is described in
detail in Ref.~\onlinecite{rftor}. In the light of the composite cylindrical resonant response model developed in Sect.~\ref{s2}, we shall
modify the resonant  response component of the aforementioned toroidal  model; this component is described in Appendix~D of Ref.~\onlinecite{rftor}.

Equations~(D1) and (D2) in Ref.~\onlinecite{rftor} are modified such that 
\begin{align}\label{e15}
\left(\hat{W}_k + \hat{\delta}_k\right){\cal S}_k\,\frac{d\hat{\mit\Psi}_k}{d\hat{t}} &= \sum_{k'=1,K}\hat{E}_{kk'}\hat{\mit\Psi}_{k'}\,\cos(\varphi_k-\varphi_{k'}-\xi_{kk'})
+\hat{E}_{kk}\,\hat{\chi}_k\,\cos(\varphi_k-\zeta_k),\\[0.5ex]
{\cal S}_k&=\frac{\tau_R(\hat{r}_k)}{\tau_A},
\end{align}
respectively. Here, 
\begin{align}
\hat{W}_k &=\frac{2\,{\cal I}}{\hat{a}\,\hat{r}_k}\left(\frac{q}{g\,s}\right)_{\hat{r}_k}^{1/2} |\hat{\mit\Psi}_k|^{\,1/2},\\[0.5ex]
\hat{\delta}_k &= \frac{\delta_{\rm SC}(\hat{r}_k)}{R_0\,\hat{a}\,\hat{r}_k},
\end{align}
where $\delta_{\rm SC}(\hat{r})$ is the semi-collisional linear layer width specified in Ref.~\onlinecite{rf1}.

@@

Equation~(D6) of Ref.~\onlinecite{rftor} is modified such that
\begin{align}\label{e19}
\left(\hat{W}_k +\hat{\delta}_k\right){\cal S}_k\left(\frac{d\varphi_k}{d\hat{t}}-\hat{\varpi}_k\right)&=
- \sum_{k'=1,K}\hat{E}_{kk'}\,\hat{\mit\Psi}_{k'}\,\sin(\varphi_k-\varphi_{k'}-\xi_{kk'})-\hat{E}_{kk}\,\hat{\chi}_k\,\sin(\varphi_k-\zeta_k),
\end{align}
where
\begin{equation}
\hat{\varpi}_k= \hat{\varpi}_{k0}+m_k\,{\mit\Delta}\hat{\mit\Omega}_\theta(\hat{r}_k,\hat{t})-n\,{\mit\Delta}\hat{\mit\Omega}_\phi(\hat{r}_k,\hat{t}),
\end{equation}
and $\hat{\varpi}_{k0}= \varpi_{k0}\,\tau_A$, $\hat{\mit\Omega}_\theta= {\mit\Delta}{\mit\Omega}_\theta\,\tau_A$, and $\hat{\mit\Omega}_\phi= {\mit\Delta}{\mit\Omega}_\phi\,\tau_A$. Here, $\varpi_{k0}$
is the so-called ``natural frequency'' at the $k$th rational surface; this quantity is defined as the helical phase velocity of a naturally unstable island chain, resonant at the surface, in the
absence of an RMP (or any other island chains). On the other hand,  ${\mit\Delta}{\mit\Omega}_\theta$ and ${\mit\Delta}{\mit\Omega}_\phi$ are the changes in the plasma
poloidal and toroidal velocity profiles, respectively, induced by the electromagnetic torques that develop in the vicinities of the various rational surfaces in the plasma.
Equation~(\ref{e19}) represents
a major improvement to Eq.~(D6) that takes into  account the finite slip-frequencies of the magnetic island chains induced at
the various rational surfaces in the plasma. In the absence of a finite slip-frequency, the right-hand side of Eq.~(\ref{e19}) would be zero.
